### 2019-09-09 submission/manuscript.Rmd
Edits from Pat

**Omeprazole treatment did not promote susceptibility to *C. difficile* infection in mice.** Next, we examined whether omeprazole treatment altered susceptibility to *C. difficile* infection in mice. After omeprazole treatment or clindamycin treatment, mice were challenged with 10^3^ *C. difficile* 630 spores. While *C. difficile* colonized the clindamycin-treated mice, it did not colonize the omeprazole-treated mice (Figure 2A). Interestingly, only 1 cage of mice that received both omeprazole and clindamycin were colonized, while the other cage of mice were resistant (Figure 2A). The greatest shifts in bacterial communities occurred in the clindamycin-treated mice (Figure 2B, S2). Regardless of whether the mice became colonized, all of the mice had cleared *C. difficile* within 5 days (Figure 2A), suggesting that omeprazole did not affect the rate of clearance. Our results suggest that omeprazole treatment had no effect on bacterial community resistance to *C. difficile* colonization in mice. Instead most of the differences between the 3 treatment groups appeared to be driven by clindamycin administration (Figure 2C, S2). These findings demonstrated that high dose omeprazole treatment did not promote susceptibility to *C. difficile* colonization.

**Conclusions.** The PPI omeprazole did not meaningfully impact the gut microbiota and did not promote *C. difficile* infection in mice. Our findings that omeprazole treatment had minimal impact on the fecal microbiome were comparable to another PPI mouse study that indicated the PPI lansoprazole had more of an effect on the small intestinal microbiota compared to the fecal microbiota [@Yasutomi2018]. The same group demonstrated lansoprazole treatment increased the stomach pH in mice [@Yasutomi2018], which may improve survival of bacteria passing through the stomach. We did not find significant changes in the relative abundances of the taxa observed to be significantly impacted by PPI use in human studies. However, 3 of the human-associated taxa were absent or at low abundance in our mice. Although our fecal microbiota findings are comparable to what has been shown in another mouse study [@Yasutomi2018], whether PPI-induced changes in specific bacterial abundances observed in humans play a role in CDIs remains to be determined.    
Although several *C. difficile* mouse model studies have shown that PPIs have an effect on CDIs with or without additional antibiotic treatment [@Kaur2007; @kaur2011effect; @hung2015proton], there were insufficient controls to attribute the effect solely to PPI treatment. One group administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice and then challenged with *C. difficile* demonstrated that PPI treatment alone resulted in detectable *C. difficile* in the stool 1 week after challenge, however there was detectable *C. difficile* in mice not treated with antibiotics [@Kaur2007; @kaur2011effect]. The other mouse study demonstrated the antibiotic/PPI esomeprazole-treated mice developed more severe CDIs compared to antibiotic-treated mice, but the researchers did not have a group treated with just esomeprazole for comparison [@hung2015proton]. We tested the same high 40 mg/kg PPI dose and expanded pre-treatment to 7 days before challenge to test the impact of omeprazole treatment alone on our CDI mouse model. Additionally, we have previously demonstrated that mice from our breeding colony are resistant to *C. difficile* 630 colonization without antibiotic treatment [@Jenior2017], ensuring there was not already partial susceptibility to *C. difficile* before treatment. The additional controls in our study allowed us to assess the contribution of omeprazole alone to *C. difficile* susceptibility in mice.  
Our study also extended previous work examining PPIs and *C. difficile* in mice by incorporating the contribution of the intestinal microbiota. We found omeprazole had no significant impact on bacterial taxa within the murine intestinal microbiota over the 16-day experiment. In contrast to previous work with PPIs [@Kaur2007; @kaur2011effect; @hung2015proton], omeprazole did not alter *C. difficile* colonization resistance in mice. 16S rRNA sequencing suggested that *Streptococcus* and *Enterococcus* are rare genera in our C57BL/6 mouse colony and variation in these 2 genera have been observed across other facilities and vendors [@Rausch2016; @Robertson2019]. These genera could be important contributors to the associations between PPIs and CDIs in humans, and could be a contributing factor to our observation that PPI treatment had no effect on *C. difficile* colonization in our CDI mouse model. Gastrointestinal physiological differences, particularly the higher stomach pH in mice (pH 3-4) compared to humans (pH 1) [@Hugenholtz2017] could also explain why omeprazole had a limited impact on the murine microbiome. The microbiota and physiological differences between humans and mice may limit the usefulness of employing mouse models to study the impact of PPIs on the microbiota and CDIs.   

>>>>

**Omeprazole treatment did not promote susceptibility to *C. difficile* infection in mice.** Next, we examined whether omeprazole treatment altered susceptibility to *C. difficile* infection in mice. After omeprazole treatment or clindamycin treatment, mice were challenged with 10^3^ *C. difficile* 630 spores. Although *C. difficile* colonized the clindamycin-treated mice, it did not colonize the omeprazole-treated mice (Figure 2A). Interestingly, only 1 cage of mice that received both omeprazole and clindamycin were colonized, while the other cage of mice were resistant (Figure 2A). The greatest shifts in bacterial communities occurred in the clindamycin-treated mice (Figure 2B, S2). Regardless of whether the mice became colonized, all of the mice had cleared *C. difficile* within 5 days (Figure 2A), suggesting that omeprazole did not affect the rate of clearance. Our results suggest that omeprazole treatment had no effect on bacterial community resistance to *C. difficile* colonization in mice. Instead most of the differences between the 3 treatment groups appeared to be driven by clindamycin administration (Figure 2C, S2). These findings demonstrated that high dose omeprazole treatment did not promote susceptibility to *C. difficile* colonization.

**Conclusions.** The PPI omeprazole did not meaningfully impact the structure of the gut microbiota and did not promote *C. difficile* infection in mice. Our findings that omeprazole treatment had minimal impact on the fecal microbiome were comparable to another PPI mouse study that indicated the PPI lansoprazole had more of an effect on the small intestinal microbiota compared to the fecal microbiota [@Yasutomi2018]. The same group demonstrated lansoprazole treatment increased the stomach pH in mice [@Yasutomi2018], which may improve survival of bacteria passing through the stomach. We did not find significant changes in the relative abundances of the taxa observed to be significantly impacted by PPI use in human studies. However, 3 of the human-associated taxa were absent or at low abundance in our mice. Although our fecal microbiota findings are comparable to what has been shown in another mouse study [@Yasutomi2018], whether PPI-induced changes in specific bacterial abundances observed in humans play a role in CDIs remains to be determined.    
Although several *C. difficile* mouse model studies have shown that PPIs have an effect on CDIs with or without additional antibiotic treatment [@Kaur2007; @kaur2011effect; @hung2015proton], there were insufficient controls to attribute the effect solely to PPI treatment. One group administered 0.5 mg/kg of the PPI lansoprazole daily for 2 weeks to mice and then challenged with *C. difficile* demonstrated that PPI treatment alone resulted in detectable *C. difficile* in the stool 1 week after challenge, however there was detectable *C. difficile* in mice not treated with antibiotics [@Kaur2007; @kaur2011effect]. The other mouse study demonstrated antibiotic/esomeprazole-treated mice developed more severe CDIs compared to antibiotic-treated mice, but the researchers did not have a group treated with just esomeprazole for comparison [@hung2015proton]. We tested the same high 40 mg/kg PPI dose and expanded pre-treatment to 7 days before challenge to test the impact of omeprazole treatment alone on our CDI mouse model. Additionally, we have previously demonstrated that mice from our breeding colony are resistant to *C. difficile* 630 colonization without antibiotic treatment [@Jenior2017], ensuring there was not already partial susceptibility to *C. difficile* before treatment. The additional controls in our study allowed us to assess the contribution of omeprazole alone to *C. difficile* susceptibility in mice.  
Our study also extended previous work examining PPIs and *C. difficile* in mice by incorporating the contribution of the intestinal microbiota. We found omeprazole had no significant impact on bacterial taxa within the murine intestinal microbiota over the 16-day experiment. In contrast to previous work with PPIs [@Kaur2007; @kaur2011effect; @hung2015proton], omeprazole did not alter *C. difficile* colonization resistance in mice. 16S rRNA sequencing suggested that *Streptococcus* and *Enterococcus* are rare genera in our C57BL/6 mouse colony. These two genera could be important contributors to the associations between PPIs and CDIs in humans, and could be a contributing factor to our observation that PPI treatment had no effect on *C. difficile* colonization in our CDI mouse model. Gastrointestinal physiological differences, particularly the higher stomach pH in mice (pH 3-4) compared to humans (pH 1) [@Hugenholtz2017] could also explain why omeprazole had a limited impact on the murine microbiome. The microbiota and physiological differences between humans and mice may limit the usefulness of employing mouse models to study the impact of PPIs on the microbiota and CDIs.   

### 2019-09-09 submission/manuscript.tex
Edits from Pat

\textbf{Omeprazole treatment did not promote susceptibility to \emph{C.
difficile infection} in mice.} Next, we examined whether omeprazole
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After omeprazole treatment or clindamycin treatment, mice were
challenged with 10\textsuperscript{3} \emph{C. difficile} 630 spores.
While \emph{C. difficile} colonized the clindamycin-treated mice, it did
not colonize the omeprazole-treated mice (Figure 2A). Interestingly,
only 1 cage of mice that received both omeprazole and clindamycin were
colonized, while the other cage of mice were resistant (Figure 2A). The
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B, S2). Regardless of whether the mice
became colonized, all of the mice had cleared \emph{C. difficile} within
5 days (Figure 2A), suggesting that omeprazole did not affect the rate
of clearance. Our results suggest that omeprazole treatment had no
effect on bacterial community resistance to \emph{C. difficile}
colonization in mice. Instead most of the differences between the 3
treatment groups appeared to be driven by clindamycin administration
(Figure 2C, S2). These findings demonstrated that high dose omeprazole
treatment did not promote susceptibility to \emph{C. difficile}
colonization.

@@

gut microbiota and did not promote \emph{C. difficile} infection in
mice. Our findings that omeprazole treatment had minimal impact on the
fecal microbiome were comparable to another PPI mouse study that
indicated the PPI lansoprazole had more of an effect on the small
intestinal microbiota compared to the fecal microbiota (12). The same
group demonstrated lansoprazole treatment increased the stomach pH in
mice (12), which may improve survival of bacteria passing through the

@@

with antibiotics (13, 14). The other mouse study demonstrated the
antibiotic/PPI esomeprazole-treated mice developed more severe CDIs
compared to antibiotic-treated mice, but the researchers did not have a
group treated with just esomeprazole for comparison (15). We tested the
same high 40 mg/kg PPI dose and expanded pre-treatment to 7 days before

@@

already partial susceptiblity to \emph{C. difficile} before treatment.

@@

contibution of omeprazole alone to \emph{C. difficile} susceptibility in
mice.\\

@@

are rare genera in our C57BL/6 mouse colony and variation in these 2
genera have been observed across other facilities and vendors (17, 18).
These genera could be important contributors to the associations between
PPIs and CDIs in humans, and could be a contributing factor to our
observation that PPI treatment had no effect on \emph{C. difficile}
colonization in our CDI mouse model. Gastrointestinal physiological
differences, particularly the higher stomach pH in mice (pH 3-4)
compared to humans (pH 1) (19) could also explain why omeprazole had a
limited impact on the murine microbiome. The microbiota and
physiological differences between humans and mice may limit the
usefulness of employing mouse models to study the impact of PPIs on the
microbiota and CDIs.\\

@@

challenging (20). Future studies are needed to determine whether age,

@@

\textbf{Animals.} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was spit between 2 cages with 2-3 mice housed per cage and male and
female mice housed separately. All animal experiments were approved by
the University of Michigan Animal Care and Use Committee (IACUC) under
protocol number PRO00006983.

@@

\textbf{16S rRNA gene sequencing.} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The 16S rRNA sequencing libary was prepared as
described previously (21). In brief, the ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard (Zymo, CA, USA) was used as a mock
community (22) and water was used as a negative control. The V4
hypervariable region of the 16S rRNA gene was amplified with Accuprime
Pfx DNA polymerase (Thermo Fisher Scientific) using previously described
custom barcoded primers (21). The 16S rRNA amplicon library was
sequenced with the MiSeq (Illumina). Amplicons were cleaned up and
normalized with the SequalPrep Normalization Plate Kit (ThermoFisher
Scientific) and pooled amplicons were quantified with the KAPA library
quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis.} mothur (v1.40.5) was used for
all sequence processing steps (23) using a previously published protocol
(21). In brief, forward and reverse reads for each sample were combined
and low-quality sequences and chimeras were removed. Duplicate sequences
were merged, before taxonomy assignment using a modified version (v16)
of the Ribosomal Database Project reference database (v11.5) with an
80\% cutoff. Operational taxonomic units (OTUs) were assigned with the
opticlust clustering algorithm using a 97\% similarity threshold. To
adjust for uneven sequencing across samples, all samples were rarefied
to 3,000 sequences, 1,000 times. PCoAs were generated based on
Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

\textbf{Statistical Analysis.} To test for differences in relative
abundances in families and genera across our 3 different treatment
groups at different timepoints (Clindamycin, Clind. + Omep., and
Omeprazole on Day -7, 0, 2, and 9) or within the Omeprazole treatment
group across 3 timepoints (Day -7, 0, and 9), we used a Kruskal-Wallis
test with a Benjamini-Hochberg correction for multiple comparisons.

\textbf{Code availability.} The code for all sequence processing and
analysis steps as well as an Rmarkdown version of this manuscript is
available at
\url{https://github.com/SchlossLab/Tomkovich_PPI_mSphere_2019}.

@@

\hypertarget{ref-Rausch2016}{}
17. \textbf{Rausch P}, \textbf{Basic M}, \textbf{Batra A},
\textbf{Bischoff SC}, \textbf{Blaut M}, \textbf{Clavel T},
\textbf{Gläsner J}, \textbf{Gopalakrishnan S}, \textbf{Grassl GA},
\textbf{Günther C}, \textbf{Haller D}, \textbf{Hirose M},
\textbf{Ibrahim S}, \textbf{Loh G}, \textbf{Mattner J}, \textbf{Nagel
S}, \textbf{Pabst O}, \textbf{Schmidt F}, \textbf{Siegmund B},
\textbf{Strowig T}, \textbf{Volynets V}, \textbf{Wirtz S},
\textbf{Zeissig S}, \textbf{Zeissig Y}, \textbf{Bleich A},
\textbf{Baines JF}. 2016. Analysis of factors contributing to variation
in the c57bl/6J fecal microbiota across german animal facilities.
International Journal of Medical Microbiology \textbf{306}:343--355.
doi:\href{https://doi.org/10.1016/j.ijmm.2016.03.004}{10.1016/j.ijmm.2016.03.004}.

\hypertarget{ref-Robertson2019}{}
18. \textbf{Robertson SJ}, \textbf{Lemire P}, \textbf{Maughan H},
\textbf{Goethel A}, \textbf{Turpin W}, \textbf{Bedrani L},
\textbf{Guttman DS}, \textbf{Croitoru K}, \textbf{Girardin SE},
\textbf{Philpott DJ}. 2019. Comparison of co-housing and littermate
methods for microbiota standardization in mouse models. Cell Reports
\textbf{27}:1910--1919.e2.
doi:\href{https://doi.org/10.1016/j.celrep.2019.04.023}{10.1016/j.celrep.2019.04.023}.

\hypertarget{ref-Hugenholtz2017}{}
19. \textbf{Hugenholtz F}, \textbf{Vos WM de}. 2017. Mouse models for
human intestinal microbiota research: A critical evaluation. Cellular
and Molecular Life Sciences \textbf{75}:149--160.
doi:\href{https://doi.org/10.1007/s00018-017-2693-8}{10.1007/s00018-017-2693-8}.

\hypertarget{ref-Eze2017}{}
20. \textbf{Eze P}, \textbf{Balsells E}, \textbf{Kyaw MH}, \textbf{Nair
H}. 2017. Risk factors for clostridium difficile infections an overview
of the evidence base and challenges in data synthesis. Journal of Global
Health \textbf{7}.
doi:\href{https://doi.org/10.7189/jogh.07.010417}{10.7189/jogh.07.010417}.

\hypertarget{ref-Kozich2013}{}
21. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Sze2019}{}
22. \textbf{Sze MA}, \textbf{Schloss PD}. 2019. The impact of DNA
polymerase and number of rounds of amplification in PCR on 16S rRNA gene
sequence data. mSphere \textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00163-19}{10.1128/msphere.00163-19}.

\hypertarget{ref-Schloss2009}{}
23. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

@@

\includegraphics{figure_1.pdf} \textbf{Figure 1. Omeprazole treatment
had minimal impact on the murine fecal microbiota.} A. Mouse experiment
timeline and logistics. The PPI omeprazole was administered throughout
the duration of the experiment. Clindamycin was administered 1 day
before \emph{C. difficile} challenge on Day 0. Stools for 16S rRNA
sequencing analysis were collected on the days that are labeled (Day -7,
-5, -3, -1, 0, 1, 2, 3, 4, 5, 7, 9). \emph{C. difficile} CFU in the
stool was quantified daily through 6 days post-infection by anaerobic
culture. B. Principal Coordinates Analaysis (PCoA) of Bray-curtis
distances from stool samples of mice in the omeprazole treatment group
during the intial 7 days of the experiment. Each color represents stool
samples from the same mouse and lines connect sequentially collected
samples. C-D. Relative abundances of families previously associated with
PPI use in humans at the start of the experiment (C) and after 7 days of
omeprazole treatment (D). Each circle represents an individual mouse.
There were no significant differences acoss treatment groups for any of
the identified families in the sequence data at day -7 (all P-values
\textgreater{} 0.448) and day 0 (all P-values \textgreater{} 0.137),
analyzed by Kruskal-Wallis test with a Benjamini-Hochberg correction for
multiple comparisons. For C-D, the grey vertical line indicates the
limit of detection.

@@

\includegraphics{figure_2.pdf} \textbf{Figure 2. Omeprazole treatment
alone does not promote CDIs in mice.} A. \emph{C. difficile} CFUs/g
stool measured each day post \emph{C. difficile} challenge for
clindamycin, clindamycin/omeprazole, and omeprazole-treated mice. Lines
represent the mean CFU/g for each treatment group while points represent
CFU/g for individual mice within each group. The black dashed line
indicates the limit of detection. B. PCoA of of Bray-curtis distances
from stool samples collected after antibiotic treatment (last 9 days of
the experiment). Transparency of the symbol corresponds to treatment
day. Symbols represent the \emph{C. difficile} colonization status of
the mice measured 2 days post-infection. Circles represent resistant
mice (\emph{C. difficile} was undetectable in stool samples), while
X-shapes represent mice that were colonized with \emph{C. difficile},
although all mice cleared \emph{C. difficile} within 5 days of
infection. Omeprazole treated fecal samples primarily cluster together
throughout the experiment. C. Genera that vary the most across treatment
groups for stool samples collected from mice 2 days post-infection. Data
were analyzed by Kruskal-Wallis test, and no P-values were significant
after Benjamini-Hochberg correction for multiple comparisons (all
P-values \textgreater{} 0.092). The grey vertical line indicates the
limit of detection.

@@

\includegraphics{figure_s1.pdf} \textbf{Figure S1. Families within
omeprazole treated mice fluctuate over time with no overall trend in
either direction.} Relative abundance over time for
\emph{Lactobacillaceae} (A) and \emph{Ruminococcaceae} (B), 2 of the
PPI-associated familes from human PPI studies across all 3 treatment
groups. Each point represents the relative abundance for an individual
mouse stool sample, while the lines represent the mean relative
abundances for each treatment group of mice. The grey horizontal lines
indicate the limit of detection.

@@

\includegraphics{figure_s2.pdf} \textbf{Figure S2. Microbiota diversity
and richness decrease with antibiotic treatment but remain relatively
constant with omeprazole treatment.} A. Boxplots of the Shannon
Diversity Index values (A) and number of observed OTUs (B) for each
group of mice over 3 timepoints (Day -7, 0, and 9). Each circle
represents the value for a stool sample from an individual mouse.

>>>>

\textbf{Omeprazole treatment did not promote susceptibility to \emph{C.
difficile} infection in mice.} Next, we examined whether omeprazole
treatment altered susceptibility to \emph{C. difficile} infection in
mice. After omeprazole treatment or clindamycin treatment, mice were
challenged with 10\textsuperscript{3} \emph{C. difficile} 630 spores.
Although \emph{C. difficile} colonized the clindamycin-treated mice, it
did not colonize the omeprazole-treated mice (Figure 2A). Interestingly,
only 1 cage of mice that received both omeprazole and clindamycin were
colonized, while the other cage of mice were resistant (Figure 2A). The
greatest shifts in bacterial communities occurred in the
clindamycin-treated mice (Figure 2B, S2). Regardless of whether the mice
became colonized, all of the mice had cleared \emph{C. difficile} within
5 days (Figure 2A), suggesting that omeprazole did not affect the rate
of clearance. Our results suggest that omeprazole treatment had no
effect on bacterial community resistance to \emph{C. difficile}
colonization in mice. Instead most of the differences between the 3
treatment groups appeared to be driven by clindamycin administration
(Figure 2C, S2). These findings demonstrated that high dose omeprazole
treatment did not promote susceptibility to \emph{C. difficile}
colonization.

@@

structure of the gut microbiota and did not promote \emph{C. difficile}
infection in mice. Our findings that omeprazole treatment had minimal
impact on the fecal microbiome were comparable to another PPI mouse
study that indicated the PPI lansoprazole had more of an effect on the
small intestinal microbiota compared to the fecal microbiota (12). The
same group demonstrated lansoprazole treatment increased the stomach pH
in mice (12), which may improve survival of bacteria passing through the

@@

with antibiotics (13, 14). The other mouse study demonstrated
antibiotic/esomeprazole-treated mice developed more severe CDIs compared
to antibiotic-treated mice, but the researchers did not have a group
treated with just esomeprazole for comparison (15). We tested the same
high 40 mg/kg PPI dose and expanded pre-treatment to 7 days before

@@

already partial susceptibility to \emph{C. difficile} before treatment.

@@

contribution of omeprazole alone to \emph{C. difficile} susceptibility
in mice.\\

@@

are rare genera in our C57BL/6 mouse colony. These two genera could be
important contributors to the associations between PPIs and CDIs in
humans, and could be a contributing factor to our observation that PPI
treatment had no effect on \emph{C. difficile} colonization in our CDI
mouse model. Gastrointestinal physiological differences, particularly
the higher stomach pH in mice (pH 3-4) compared to humans (pH 1) (17)
could also explain why omeprazole had a limited impact on the murine
microbiome. The microbiota and physiological differences between humans
and mice may limit the usefulness of employing mouse models to study the
impact of PPIs on the microbiota and CDIs.\\

@@

challenging (18). Future studies are needed to determine whether age,

@@

\textbf{Animals.} All mouse experiments were performed with 7- to
12-week-old C57BL/6 male and female mice. Each experimental group of
mice was split between 2 cages with 2-3 mice housed per cage and male
and female mice housed separately. All animal experiments were approved
by the University of Michigan Animal Care and Use Committee (IACUC)
under protocol number PRO00006983.

@@

\textbf{16S rRNA gene sequencing.} DNA for 16S rRNA gene sequencing was
extracted from 10-50 mg fecal pellet from each mouse using the DNeasy
Powersoil HTP 96 Kit (Qiagen) and an EpMotion 5075 automated pipetting
system (Eppendorf). The 16S rRNA sequencing library was prepared as
described previously (19). In brief, the ZymoBIOMICS\textsuperscript{TM}
Microbial Community DNA Standard (Zymo, CA, USA) was used as a mock
community (20) and water was used as a negative control. The V4
hypervariable region of the 16S rRNA gene was amplified with Accuprime
Pfx DNA polymerase (Thermo Fisher Scientific) using previously described
custom barcoded primers (19). The 16S rRNA amplicon library was
sequenced with the MiSeq (Illumina). Amplicons were cleaned up and
normalized with the SequalPrep Normalization Plate Kit (ThermoFisher
Scientific) and pooled amplicons were quantified with the KAPA library
quantification kit (KAPA Biosystems).

\textbf{16S rRNA gene sequence analysis.} mothur (v1.40.5) was used for
all sequence processing steps (21) using a previously published protocol
(19). In brief, forward and reverse reads for each sample were combined
and low-quality sequences and chimeras were removed. Duplicate sequences
were merged, before taxonomy assignment using a modified version (v16)
of the Ribosomal Database Project reference database (v11.5) with an
80\% cutoff. Operational taxonomic units (OTUs) were assigned with the
opticlust clustering algorithm using a 97\% similarity threshold. To
adjust for uneven sequencing across samples, all samples were rarefied
to 3,000 sequences, 1,000 times. PCoAs were generated based on
Bray-Curtis distance. R (v.3.5.1) was used to generate figures and
perform statistical analysis.

\textbf{Statistical Analysis.} To test for differences in relative
abundances in families and genera across our 3 different treatment
groups at different timepoints (Clindamycin, Clindamycin + Omeprazole,
and Omeprazole on Day -7, 0, 2, and 9) or within the Omeprazole
treatment group across 3 timepoints (Day -7, 0, and 9), we used a
Kruskal-Wallis test with a Benjamini-Hochberg correction for multiple
comparisons.

\textbf{Code availability.} The code for all sequence processing and
analysis steps as well as a Rmarkdown version of this manuscript is
available at
\url{https://github.com/SchlossLab/Tomkovich_PPI_mSphere_2019}.

@@

\hypertarget{ref-Hugenholtz2017}{}
17. \textbf{Hugenholtz F}, \textbf{Vos WM de}. 2017. Mouse models for
human intestinal microbiota research: A critical evaluation. Cellular
and Molecular Life Sciences \textbf{75}:149--160.
doi:\href{https://doi.org/10.1007/s00018-017-2693-8}{10.1007/s00018-017-2693-8}.

\hypertarget{ref-Eze2017}{}
18. \textbf{Eze P}, \textbf{Balsells E}, \textbf{Kyaw MH}, \textbf{Nair
H}. 2017. Risk factors for clostridium difficile infections an overview
of the evidence base and challenges in data synthesis. Journal of Global
Health \textbf{7}.
doi:\href{https://doi.org/10.7189/jogh.07.010417}{10.7189/jogh.07.010417}.

\hypertarget{ref-Kozich2013}{}
19. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq illumina sequencing platform.
Applied and Environmental Microbiology \textbf{79}:5112--5120.
doi:\href{https://doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}.

\hypertarget{ref-Sze2019}{}
20. \textbf{Sze MA}, \textbf{Schloss PD}. 2019. The impact of DNA
polymerase and number of rounds of amplification in PCR on 16S rRNA gene
sequence data. mSphere \textbf{4}.
doi:\href{https://doi.org/10.1128/msphere.00163-19}{10.1128/msphere.00163-19}.

\hypertarget{ref-Schloss2009}{}
21. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

@@

\includegraphics{figure_1.pdf} \textbf{Figure 1. Omeprazole treatment
had minimal impact on the murine fecal microbiota.} A. Mouse experiment
timeline and logistics. The PPI omeprazole was administered throughout
the duration of the experiment. Clindamycin was administered 1 day
before \emph{C. difficile} challenge on Day 0. Stools for 16S rRNA
sequencing analysis were collected on the days that are labeled (Day -7,
-5, -3, -1, 0, 1, 2, 3, 4, 5, 7, 9). \emph{C. difficile} CFU in the
stool was quantified daily through 6 days post-infection by anaerobic
culture. B. Principal Coordinates Analysis (PCoA) of Bray-Curtis
distances from stool samples of mice in the omeprazole treatment group
during the initial 7 days of the experiment. Each color represents stool
samples from the same mouse and lines connect sequentially collected
samples. C-D. Relative abundances of families previously associated with
PPI use in humans at the start of the experiment (C) and after 7 days of
omeprazole treatment (D). Each circle represents an individual mouse.
There were no significant differences across treatment groups for any of
the identified families in the sequence data at day -7 (all P-values
\textgreater{} 0.448) and day 0 (all P-values \textgreater{} 0.137),
analyzed by Kruskal-Wallis test with a Benjamini-Hochberg correction for
multiple comparisons. For C-D, the grey vertical line indicates the
limit of detection.

@@

\includegraphics{figure_2.pdf} \textbf{Figure 2. Omeprazole treatment
alone does not promote CDIs in mice.} A. \emph{C. difficile} CFUs/g
stool measured each day post \emph{C. difficile} challenge for
clindamycin, clindamycin/omeprazole, and omeprazole-treated mice. Lines
represent the mean CFU/g for each treatment group while points represent
CFU/g for individual mice within each group. The black dashed line
indicates the limit of detection. B. PCoA of of Bray-Curtis distances
from stool samples collected after antibiotic treatment (last 9 days of
the experiment). Transparency of the symbol corresponds to treatment
day. Symbols represent the \emph{C. difficile} colonization status of
the mice measured 2 days post-infection. Circles represent resistant
mice (\emph{C. difficile} was undetectable in stool samples), while
X-shapes represent mice that were colonized with \emph{C. difficile},
although all mice cleared \emph{C. difficile} within 5 days of
infection. Omeprazole treated fecal samples primarily cluster together
throughout the experiment. C. Genera that vary the most across treatment
groups for stool samples collected from mice 2 days post-infection. Data
were analyzed by Kruskal-Wallis test, and no P-values were significant
after Benjamini-Hochberg correction for multiple comparisons (all
P-values \textgreater{} 0.092). The grey vertical line indicates the
limit of detection.

@@

\includegraphics{figure_s1.pdf} \textbf{Figure S1. Families within
omeprazole treated mice fluctuate over time with no overall trend in
either direction.} Relative abundance over time for
\emph{Lactobacillaceae} (A) and \emph{Ruminococcaceae} (B), 2 of the
PPI-associated families from human PPI studies across all 3 treatment
groups. Each point represents the relative abundance for an individual
mouse stool sample, while the lines represent the mean relative
abundances for each treatment group of mice. The grey horizontal lines
indicate the limit of detection.

@@

\includegraphics{figure_s2.pdf} \textbf{Figure S2. Microbiota diversity
and richness decrease with antibiotic treatment but remain relatively
constant with omeprazole treatment.} Boxplots of the Shannon Diversity
Index values (A) and number of observed OTUs (B) for each group of mice
over 3 timepoints (Day -7, 0, and 9). Each circle represents the value
for a stool sample from an individual mouse.
